---
title: "4740 Project"
author: "Chris VanKerkhove"
date: "5/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A Music Assignment Model

In this Markdown, I will document my process of creating and validating different models (with increasing complexity) for classifying music snippets into different musical genres. 

The data is taken from Kaggle from the page <https://www.kaggle.com/insiyeah/musicfeatures>. The data was generated using 30 second MP3 snippets of songs and then extracting features of the sound-bite to act as different predictors. 


```{r Setup, include = FALSE}
setwd("/Users/cvankerkhove/Documents/GitHub /Music-Classification-Model")
library(readr)
library(tibble)
library(dplyr)
library(tidyr)
library(ggplot2)
source("Helpful_Functions.R")
```

### Data Preproccessing

Checking for missing values, casting types for classification, ensuring clean data.frame

```{r Data, include = FALSE}
####Data Cleaning and Manipulation###
music.dat <- read_csv("data.csv") %>%
  tibble()
#missing values
missing <- c()
classes <- c()
for (col.name in colnames(music.dat)) {
  missing[col.name] <- sum(is.na(music.dat[[col.name]]))
  classes[col.name] <- class(music.dat[[col.name]])
}
```

```{r Data Results, echo = TRUE}
#Dataset is complete, No missing data
sum(missing)
#removing file name column ('label' column acts as dependent var)
df <- select(music.dat, -filename)
df$label <- as.factor(df$label)
tibble(df)
```

## K-Nearest-Neighbor Algorithm

### Multinomial Classification

In the following cells I will perform 10-fold CV 10 times, on the KNN algorithm classification for different values of K (number of nearest neighbors to consider during algorithm), and then plot to determine the best value of K.

```{r KNN-CV, echo=FALSE}
#running multinomial classification for different values of K
#computing error through 10-fold CV 10 times
k.values <- seq(1,30,1)
cv.errors <- c()
for (i in 1:length(k.values)) {
  #errors from 10-fold, 10 iteration CV
  out2 <- run.knn.cv(df, k.values[i], levels(df$label), 10, 10)
  cv.errors[i] <- out2
}
plot(k.values, cv.errors)
```

Based on the plot from this validation method, for reducing overall error (across all classification classes), the optimal **K= 5**. 
Next, I will run 20 iterations of KNN using the validation set approach (60% training), this time recording and plotting the error of each respective genre

```{r Genre-validation set error, include=FALSE}
#running 10 iterations of validation set knn
#computing error classification error by each class
genres <- levels(df$label)
#dataframe to store error data for each genre across iterations
errors.df <- data.frame(matrix(ncol=10, nrow=0))
colnames(errors.df) <- genres
#number of iterations to average error over
n <- 20

for (itr in 1:n) {
  out <- run.knn(df, 5, levels(df$label), 0.6)
  d <- as.data.frame(out[[1]])
  d1 <- as.data.frame(out[[1]]) %>%
    group_by(test.label) %>%
    summarise(tot = sum(Freq)) %>%
    ungroup()
  
  for (i in 1:length(genres)){
    numer <- filter(d, d$knn.pred == genres[i] & d$test.label == genres[i])$Freq
    denom <- filter(d1, d1$test.label == genres[i])$tot
    errors.df[itr, genres[i]] <- 1 - (numer/denom)
  }
}
errors <- summarise_each(errors.df, list(.=mean))
```
Errors:
```{r Errors Plot , echo=FALSE}
##ggplot2 barplotting
#getting error data in a valid input to ggplot2
errors.num <- c()
for (i in 1:ncol(errors)) {
  errors.num[genres[i]] <- errors[[i]]
}

errors.df <- cbind(genre = rownames(as.data.frame(errors.num)), as.data.frame(errors.num))
rownames(errors.df) <- 1:nrow(errors.df)
#plotting
ggplot(errors.df, aes(genre, errors.num)) +
  geom_bar(stat = "identity")
```

### Binary Classification

Next, I will use the KNN algorithm, but as a binary classifier, i.e. for each genre I will update the dataframe and change every other genre to "other"
so that the column as a factor only has 2 levels (said genre, and other). The goal of this is to improve the accuracy of each genre without multinomial classification.
Furthermore, I will use 10 fold CV 10 times f choose a seperate optimal K value for each genre; providing those values of K and plotting the errors. 

Another assumption, I randomly select 100 "other" data points to keep the dataset balanced, as we will only have 100 points for each individual genre

``` {r KNN Binary Classification, include=FALSE}
#binary classification for each genre
#for each genre
#take data points from genre (100)
#select 100 random data points from non-genre and label as 'other'
#run KNN 
genres <- levels(df$label)
#list for recording errors of KNN
errors <- c()
#lists for finding best k-value for each genre
k.values <- seq(1,20,1)
best.k <- c()

for (i in 1:length(genres)) {
  print(genres[i])
  df.genre <- filter(df, !(label == genres[i])) %>%
    sample_n(100, replace = FALSE) %>%
    mutate(label = 'other') %>%
    bind_rows(filter(df, label == genres[i]))
  df.genre$label <- as.factor(df.genre$label)
  #out <- run.knn(df.genre, 3, levels(df.genre$label), 0.5)
  #error[genres[i]] <- out[[2]]
  #tables[[genres[i]]] <- out[[1]]
  #list for storing error values of each K 
  k.list <- c()
  for (k in 1:length(k.values)) {
    out <- run.knn.cv(df.genre, k.values[k], levels(df.genre$label), 10, 10)
    k.list[k] <- out
  }
  best.k[genres[i]] <- which.min(k.list)
  errors[genres[i]] <- min(k.list)
}
```

```{r Plotting Error Values, echo=FALSE}
##ggplot2 barplotting
#getting error data in a valid input to ggplot2
errors.df <- cbind(genre = rownames(as.data.frame(errors)), as.data.frame(errors),
                        as.data.frame(best.k))
rownames(errors.df) <- 1:nrow(errors.df)
errors.df$best.k <- as.factor(errors.df$best.k)

ggplot(errors.df, aes(genre, errors, fill=best.k)) +
  geom_bar(stat = "identity")
```


## Decision Trees and Extenstions
In the next section, I will attempt to classify the data using decision tree algorithms, and various extensions
such as pruning the tree, random forests, and boosting
### Pruned Decision Tree

In the following cell I use a single decision tree (pruned via 10 fold cv for best k parameter) to perform multinomial classification
and plot the classification errors across genres

```{r Pruned Tree Multinomial Classification, include = FALSE}
####Decision Trees####
library(tree)
#variable name 'df' throws error in prune.misclass() function

df.tree <- df
tree.music <- tree(label~., df.tree)
cv1 <- cv.tree(tree.music, FUN = prune.misclass)
#validation set
train <- sample(1:1000, 700, replace=FALSE)
tree.music <- tree(label~., df.tree, subset=train)
#using best k parameter from cv
pruned.tree.music <- prune.misclass(tree.music, best = cv1$size[which.min(cv1$dev)])
tree.pred <- predict(pruned.tree.music, df.tree[-train,], type='class')
t <- table(tree.pred, df.tree[-train,]$label )
#Getting each genre classification error
genres <- levels(df.tree$label)
errors <- c()
for (i in 1:length(genres)) {
  num <- t[genres[i], genres[i]]
  denom <- 0
  for (i2 in 1:nrow(t)) {
    denom <- denom + t[genres[i], i2]
  }
  errors[genres[i]] <- 1 - (num / denom)
}
```
##### Tree Summary
```{r Tree summary, echo = FALSE}
summary(pruned.tree.music)
```
##### Multinomial Confusion Matrix
```{r Confusion Matrix 1, echo = FALSE}
t
```


```{r Plotting Tree Multinomial Errors, echo = FALSE}
errors <- replace_na(errors, 1)
#ggplot2 barplot
errors.df <- cbind(genre = rownames(as.data.frame(errors)), as.data.frame(errors))
rownames(errors.df) <- 1:nrow(errors.df)

ggplot(errors.df, aes(genre, errors)) +
  geom_bar(stat = "identity")
```

In the next cell I will perform the same process, but only using sub-setted data for binary classification again (200 rows)
pruning each tree by their best k parameter using cross-validation

```{r Pruned Tree binomial classification, echo = FALSE}
genres <- levels(df.tree$label)
errors <- c()
for (i in 1:length(genres)) {
  #modifying main dataset to be a balanced 200 rows each time
  df.genre <- filter(df.tree, !(label == genres[i])) %>%
    sample_n(100, replace = FALSE) %>%
    mutate(label = 'other') %>%
    bind_rows(filter(df, label == genres[i]))
  df.genre$label <- as.factor(df.genre$label)
  #Creating a tree and pruned tree using training data
  train <- sample(1:nrow(df.genre), nrow(df.genre)*0.70)
  tree.1genre <- tree(label~., df.genre[train,])
  cv1 <- cv.tree(tree.1genre, FUN = prune.misclass)
  pruned.tree <- prune.misclass(tree.1genre, best=cv1$size[which.min(cv1$dev)])
  #predictions train
  pred <- predict(pruned.tree, newdata=df.genre[-train,], type='class')
  t <-table(pred, df.genre[-train,]$label)
  errors[genres[i]] <-  1 - mean(pred == df.genre[-train,]$label)
}
```

```{r Plotting Tree Binomial Errors, echo = FALSE}
errors <- replace_na(errors, 1)
#ggplot2 barplot
errors.df <- cbind(genre = rownames(as.data.frame(errors)), as.data.frame(errors))
rownames(errors.df) <- 1:nrow(errors.df)

ggplot(errors.df, aes(genre, errors)) +
  geom_bar(stat = "identity")
```